Viscous Erosion in Porous Media

### 2018-03-07 docs/methods/erosion_methods.tex
small changes to paper

A challenge with erosion is that it creates corners at points on the
body where the shear stress vanishes.  While recent work does allow for
integral equation methods to resolve corners, we employ a different
method.  
\todo[inline]{Maybe point out some drawbacks of using one of these BIE methods that allows corners. For example, are these methods more conceptually complex and/or more expensive? Maybe they are not quite at a mature stage for application? Another point: in the simulations we cannot predict where corners will form and so would have to identify them on the fly. Would this cause a problem if we tried to use a corner-resolving BIE?}
\todo[inline]{Could say our approach is simpler because  by smoothing corners, we can use the bare-bones BIE methods that are more well established compared to these newer, corner-resolving methods. Just some thoughts.}
First, we add a regularization term to the shear stress to
penalize regions of high curvature.  Second, we apply a Gaussian filter
to smooth the absolute value of the shear stress before the body shape
is updated.

@@

\todo[inline]{These are some of my thoughts on reworking this first paragraph. See if you like it.}
We begin by making two fundamental design choices: first, we restrict ourselves to two spatial dimensions, and second the bodies are considered immobile (i.e.~fixed to their initial locations). The restriction to 2D allows us to simulate the erosion of many bodies (order 50-100) with high resolution, giving us the ability to accurately characterize the erosion of a porous medium. While BIE methods are available in 3D, the extra computational cost limits the number of bodies that could be practically simulated, making porous-medium simulations less feasible. The second assumption of immobile bodies affords a simple framework to study the effects of erosion in isolation. While other effects, such as sedimentation or transport, undoubtedly play a role in real applications, these effects would introduce additional complexity, both conceptually and numerically.

\todo[inline]{This is the original below:}
There are two main limitations of the current
work.  First, the flow is assumed to be two-dimensional.  However, by
assuming that the flow is two-dimensional, we are able to simulate the
erosion of more bodies and with higher resolution than would be
available in three dimensions.  Moreover, several flows can be described
to good accuracy under a two-dimensional approximation, but certain
phenomena occur only in three dimensions.  Extending the current work to
three dimensions is possible with a boundary integral equation
formulation.  In particular, the fluid solvers exist~\cite{mal-bir2015},
and the main challenges of a BIE solver have been addressed by other
groups including surface representation~\cite{yin-bir-zor2006,
vee-rah-bir-zor2011}, near-singular integration~\cite{yin-bir-zor2006,
kli-tor2016b}, and fast summation methods~\cite{yin-bir-zor2004,
ros-ols2015}.  The more significant challenge of extending our work to
three dimensions is guaranteeing that the mesh does not become
distorted, which can be done, for example, by using an extension of the
two-dimensional {\thL} formulation~\cite{amb-sie-tlu2013}.

The second limitation is that the eroding bodies are fixed.  Mitchell
and Spagnolie~\cite{mit-spa2016} simulate a single mobile eroding body
in three dimensions~\cite{mit-spa2016}, but additional challenges are
caused when there are multiple bodies. In particular, especially in
dense suspensions, bodies can become extremely close to one another, and
the interaction between the bodies is difficult to resolve unless the
underlying mesh is refined to levels that are intractable. In future
work we plan to consider mobile eroding bodies by including a small
force and torque to prevent unphysical contact~\cite{lu-rah-zor2017}.

@@

\todo[inline]{Dump of text}
Other models for erosion include ones that use an erosion rate that is
proportional to a power of the magnitude of the shear
stress~\cite{par-izu2000}, or ones that use a critical shear stress
below which erosion does not occur~\cite{lag2000}.
\todo[inline]{Dump of text}

@@

We use a quasi-static approximation.  That is, at each time step, we
freeze the geometry, solve the incompressible Stokes equations, and then
update the geometry according to the erosion model.  The two steps
inside the red rectangle describes a time step in the quasi-static
approximation.
\todo[inline]{Appeal to separation of scales to justify this quasi-static approximation; i.e.~the timescale of erosion is much longer than that of the flow.}

%A flow diagram is illustrated in Figure~\ref{fig:workflow}.  
%\begin{figure}[htpb]
%  \centering
%  \input{figs/workflow.tikz}
%  \caption{\label{fig:workflow}A caption.}
%\end{figure}

@@

>>>>

A challenge with erosion is that it creates corners at points on the
body where the shear stress vanishes.  There is recent work for solving
boundary integral equations in geometries with corners by using
carefully chosen local expansions around each corner~\cite{rac-ser2017,
ser-rok2016} or panel-based graded meshes~\cite{hel2011,
gil-hao-mar2014, bre2012}.  The limitation of these methods is that they
require the corner's location, a non-uniform spaced mesh, and some of
the methods have a large condition number.  So that we can use a uniform
grid, we instead slightly modify the geometry to maintain sufficient
regularity.  First, we add a regularization term to the shear stress to
penalize regions of high curvature.  Second, we apply a Gaussian filter
to smooth the absolute value of the shear stress before the body shape
is updated.

@@

%\todo[inline]{These are some of my thoughts on reworking this first paragraph. See if you like it.}
We begin by making two fundamental design choices: first, we restrict ourselves to two spatial dimensions, and second the bodies are considered immobile (i.e.~fixed to their initial locations). The restriction to 2D allows us to simulate the erosion of many bodies (order 50-100) with high resolution, giving us the ability to accurately characterize the erosion of a porous medium. While BIE methods are available in 3D, the extra computational cost limits the number of bodies that could be practically simulated, making porous-medium simulations less feasible. The second assumption of immobile bodies affords a simple framework to study the effects of erosion in isolation. While other effects, such as sedimentation or transport, undoubtedly play a role in real applications, these effects would introduce additional complexity, both conceptually and numerically.

%\todo[inline]{This is the original below:}
%There are two main limitations of the current
%work.  First, the flow is assumed to be two-dimensional.  However, by
%assuming that the flow is two-dimensional, we are able to simulate the
%erosion of more bodies and with higher resolution than would be
%available in three dimensions.  Moreover, several flows can be described
%to good accuracy under a two-dimensional approximation, but certain
%phenomena occur only in three dimensions.  Extending the current work to
%three dimensions is possible with a boundary integral equation
%formulation.  In particular, the fluid solvers exist~\cite{mal-bir2015},
%and the main challenges of a BIE solver have been addressed by other
%groups including surface representation~\cite{yin-bir-zor2006,
%vee-rah-bir-zor2011}, near-singular integration~\cite{yin-bir-zor2006,
%kli-tor2016b}, and fast summation methods~\cite{yin-bir-zor2004,
%ros-ols2015}.  The more significant challenge of extending our work to
%three dimensions is guaranteeing that the mesh does not become
%distorted, which can be done, for example, by using an extension of the
%two-dimensional {\thL} formulation~\cite{amb-sie-tlu2013}.
%
%The second limitation is that the eroding bodies are fixed.  Mitchell
%and Spagnolie~\cite{mit-spa2016} simulate a single mobile eroding body
%in three dimensions~\cite{mit-spa2016}, but additional challenges are
%caused when there are multiple bodies. In particular, especially in
%dense suspensions, bodies can become extremely close to one another, and
%the interaction between the bodies is difficult to resolve unless the
%underlying mesh is refined to levels that are intractable. In future
%work we plan to consider mobile eroding bodies by including a small
%force and torque to prevent unphysical contact~\cite{lu-rah-zor2017}.

@@

@@

To perform the simulations, we freeze the geometry at each time step,
solve the incompressible Stokes equations, and then update the geometry
according to the erosion model.  The separation between these two
processes is justified since the time scale of erosion is much longer
than that of the flow.

%We use a quasi-static approximation.  That is, at each time step, we
%freeze the geometry, solve the incompressible Stokes equations, and then
%update the geometry according to the erosion model.  The two steps
%inside the red rectangle describes a time step in the quasi-static
%approximation.
%\todo[inline]{Appeal to separation of scales to justify this quasi-static approximation; i.e.~the timescale of erosion is much longer than that of the flow.}

@@

Other models for erosion include ones that use an erosion rate that is
proportional to a power of the magnitude of the shear
stress~\cite{par-izu2000}, or ones that use a critical shear stress
below which erosion does not occur~\cite{lag2000}.
